---
title: "Paper Information Statistics"
author: "Peiyuan"
format: gfm
editor: visual
---

## Project Introduction

This project extracts structural statistics from academic paper PDFs: page
count, abstract word count, section word counts, figure/table counts, caption
density, reference count, and review diagnostics.

The current implementation is designed as a configurable pipeline rather than
a script tied to one or two journals. Journal-specific rules live in
`R/profiles.R`; reusable parsing and summarisation functions live in
`R/analyze_paper.R`.

## Quick Start

Requirements:

- R >= 4.1.0
- R packages: `pdftools`, `purrr`, `dplyr`, `tidyr`, `readr`, `stringr`,
  `tibble`, `here`

1. Restore or install the required R packages.

```{r}
#| eval: false
renv::restore()
```

2. Put PDF files in one or more input folders.

3. Edit `default_input_dirs` in `paper_information_statistics.R`, or set the
   `paperstats.input_dirs` option, if you want to analyze different folders.

4. Run the script.

```{r}
#| eval: false
source("paper_information_statistics.R")
```

To run a one-off folder without editing the script:

```{r}
#| eval: false
options(paperstats.input_dirs = c("path/to/pdf_folder"))
source("paper_information_statistics.R")
```

## Outputs

The script writes five CSV files to `output/`:

- `paper_metrics.csv`: one row per paper, including section counts and detected
  structural features. `IsStandardStructure` means required sections were
  found; `NeedsReview` means the paper has missing sections, parsing errors, or
  warning-level signals such as unusually long sections.
- `summary_metrics_by_journal.csv`: aggregate statistics by journal for papers
  with standard required structure.
- `summary_metrics_overall.csv`: aggregate statistics across all journals.
- `diagnostics.csv`: papers with `NeedsReview = TRUE`.
- `heading_candidates.csv`: probable heading lines extracted from papers with
  `NeedsReview = TRUE`, useful for extending profiles.

## Adding a New Journal

Add a profile to `paper_profiles` in `R/profiles.R`.

```{r}
#| eval: false
paper_profiles$MyJournal <- make_profile(
  name = "MyJournal",
  header_patterns = c(
    "My Journal Name",
    "^\\s*https?://doi\\.org/"
  )
)
```

Then add the PDF folder to `default_input_dirs` in
`paper_information_statistics.R`, or pass it through `paperstats.input_dirs`.
If the folder name matches the profile name, that profile is used automatically.
If no matching profile exists, the default profile is used.

## Current Approach

The pipeline is intentionally rule-first:

- clean page-level noise such as page numbers, journal headers, and DOI lines;
- detect common paper sections with configurable regular expressions;
- map section variants such as `Materials and Methods` or `Background` to
  standard fields;
- record diagnostics for low-confidence or non-standard papers.

This keeps the system transparent and easy to debug. Machine learning or deep
learning can be added later for OCR, layout detection, caption boundary
detection, or semantic section classification once enough failure cases have
been collected.
